---
title: "Reading Solar Data -- Part 1"
author: "Veda Nayak"
date: "2025-04-01"
output: 
  html_document:
    number_sections: true
    highlight: tango
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("C:/cygwin64/home/vedan/Code/sta141b/assignment1code.R")

```

# Overview 

## What to Submit
 • A description of your approach and the important steps.
 • The file(s) with the R functions and the code to use those functions to read the data.
 • A description of how you verified the results were correct with numerical summaries, tables, plots that “prove” this is
 correct, along with a description of any manual verification you did.
 • The R code you used to verify the results

## Objective: 
For this assignment, you are to write functions that will read the data in the .clm (climate) file in each zip file into a data.frame. The resulting *data.frame should have 9 columns– the 6 columns in the data set and also the day, month and hour of each observation.*
Write functions (rather than one or more R commands) to read each of these tables, as you need to apply these to the contents of the 5 ZIP files. Also, you will most likely need to run the code multiple times to iteratively modify it and verify it is correct. Use functions rather than repeating the same code for each file and table

## Verification: Describe the approaches / processes taken to verify your findings
Some ideas for verification:
 • manually comparing individual values in the files and the results,
 • computing summary statistics from the results, and
 • visualizing the results
 • programmatically verifying the results

# Trial Run on a Small Dataset (three days)

I created a dataset taking lines 1 - 137 of the "USA_CA_Mammoth.Yosemite.AP.723894_TMYx.2009-2023.clm" file so that I could work with a smaller dataset. 

In addition to the columns given in the prompt, I included a column for the exact date (day_month) and for the location to uniquely identify each row when working with the dataframes later.

To make one table from one file, I created the following functions (some of which were adapted from lecture):

 • *fileName*: Make the complete filepath name from the filename
 • *getBlocks*: Get the line blocks that correspond to the data for one day
      - Get grouped start and ending line indices for the blocks. The start of the block is found by searching for "* day ". The end of the line is found by adding 24 to that value
      - The cumsum function is applied to the logical or of the start and stop pairs to get the line blocks
 • *getTables*: 
 • *makeTables*:
 • *readTableData*: 
 • *splitDataLines*: 

```{r}
table <- makeTables(fn)
head(table)
  
```

# Generating dataframes from the actual dataset

```{r echo=FALSE}
setwd("C:/cygwin64/home/vedan/Code/sta141b/Solar1/unzip")
files <- list.files(path="C:/cygwin64/home/vedan/Code/sta141b/Solar1/unzip", pattern="clm")
files <- lapply(files, fileName)
df_list = lapply(files, makeTables)
df_merged <- mergeTables(df_list)

df = as.data.frame(df_list[1])
    colnames(df) <- dfNames
    
    for (i in 2:length(df_list)){
      
      addition <- as.data.frame(df_list[i])
      colnames(addition) <- dfNames
      df <- bind_rows(df, addition)
      
    }

df

head(df_merged)
tail(df_merged)
```


## Validating the Dataframe

I did several things to validate my dataframe was correct:
• Determine if we have an expected number of rows and columns
• Check if the temperature rises in the day and falls in the night

### Number of Rows and Columns
Since we have 5 locations, 365 days per location, and 24 hours per day, the total number of rows we would expect is 5 * 365 * 24 = 43,800 rows

### Graph the data across the hour
I made a function to 

```{r}

tempByDayScatter <-  ggplot(df_merged, aes(x = hour, y = dry_bulb_temp, color = location)) +
      geom_point() +
      labs(title = "Dry Bulb Temperature vs. Hours in a Day",
           xlab = "Hour",
           ylab = "Dry Bulb Temperature",
           fill = "Date") +
      facet_wrap(~ day_month)

```
FINISH

# Acknowledgments

[Extracting the filename from the file path in the 'makeTables' function]("https://stackoverflow.com/questions/2548815/find-file-name-from-full-file-path")

# Discussion in Class
- In this case you could unzip them manually if you want, it's all chill.
- For verificiation you can also look to see if there are expected missing values
- Look at location / elevation + plot it accordingly if you would like
- How did I extract the blocks of data + how did I read each one
- Possibly make a bunch of small functions hm
- Explain iteration process (bottom up process)
- Check each column is the right class

```{r}

# vector for starts and stops
starts = which(w)
ends = which(w2)

for (i in 1:length(start)){
  
  ll[starts[i]:ends[i]]
  
}

# OR 
z = mapply(function(i,j), starts, ends, SIMPLIFY = TRUE) # a little bit more general than a forloop
# mapply takes two vectors and returns f(x, y)
# lapply takes a vector, applies a function, and returns a vector

# OR

# compare the two vectors, if either w or w2 is true return true
w3 = w | w2
which(w3) # --> list of starts and stops [5, 112, 2323, 2430], expected
cumsum(w3) # cumulative sum, work on entire vectors, adds numbers [ False --> 0, True --> 1]
# 0 0 0 0 1 [inside table you keep adding 0 to 1] 1 1 1 ... 2 [ end of alignment] ... 3[next table] ... 4 [ end of next table ]

# 'These lines are in group 1, these are in group 2...' just group the lines by number

g = split(ll, cumsum(w3)) # give me categories for each element
# easier than counting indices in tables

# odd numbers will be the tables we want

# NOW WE'VE EXTRACTED THE LINES FOR EACH TABLE

tbl = g[[2]]
con = textConnection(tbl[-(1:4)]) # treat my text as a file please



```

Old
I treated the .clm file like a .txt file FINISH
I realized that since the data spans across several years, "day 1 month 1" repeats in the data. If we want to look at trends across data, or if we want to look at a particular day, we would not be able to do so since we do not have a unique identifier. To remedy this, I first made a column for the year, which increases if the previous month has a larger value than the current month (ex. previous month = 11 and current month = 1). By not making the condition specific to months 12 and 1, it allows for missing monthly data. 
After determining the year, I created an column for the exact date in day-month-year format.

```{r}
# Get the line number that starts each day
# https://www.spsanderson.com/steveondata/posts/2024-09-03/
matching_indices <- grep("* day  ", lines.list)
print(matching_indices)
print(lines.list[matching_indices])

test_frame <- as.data.frame(matrix(0, nrow = 24*length(matching_indices)-1, ncol = 10))
colnames(test_frame) <- c('day', 'month', 'hour', 'diffuse_solar_horizontal', 'dry_bulb_temp', 'normal_solar_intensity', 'wind_speed', 'wind_direction', 'rel_humidity', 'year')

start = 1
previous_month = 12
year = 2008

for (i in 1:length(matching_indices)){
  
  # get the text contents for a particular day
  line_index = as.integer(matching_indices[i])
  day_info <- lines.list[line_index:(line_index+24)]
  day_split <- strsplit(day_info[1], split = " ")
  unlist(day_split)
  
  # day and month information
  day <- as.integer(day_split[[1]][4])
  month <- as.integer(day_split[[1]][7])
  
  if (previous_month > month){year = year + 1}
  
  # fill in the hour and remaining 6 columns
  if (i == 1){
    start = 1
    stop = 24
    }
  else {
    start = (i - 1) * 24 + 1
    stop = start + 23
    }
  
  hour = 0
  
  # remove the first line
  line_index = matching_indices[i]
  index_plus_one = line_index + 1
  ending_index = line_index + 24
  data_info <- lines.list[index_plus_one:ending_index]
  
  line_index = 1
  for (j in start:stop){
    
    test_frame[j,1] <- day
    test_frame[j,2] <- month
    test_frame[j,10] <- year
   
    if (hour == 24){break}
    
    # add in the hours 
    hour = hour + 1
    
    test_frame[j,3] <- hour
    
    line = strsplit(data_info[line_index], split = ',')

    test_frame[j, 4] <- line[[1]][1]
    test_frame[j, 5] <- line[[1]][2]
    test_frame[j, 6] <- line[[1]][3]
    test_frame[j, 7] <- line[[1]][4]
    test_frame[j, 8] <- line[[1]][5]
    test_frame[j, 9] <- line[[1]][6]
  
    line_index = line_index + 1
    
    }
    
    previous_month = month
  }
  
test_frame$exact_date <- interaction(test_frame$day, test_frame$month, test_frame$year, sep = '-')
test_frame$location <- tools::file_path_sans_ext(fn)
# https://stackoverflow.com/questions/29113973/get-filename-without-extension-in-r
=
```

