---
title: "Assignment2_ReadingServerAlertData"
author: "Veda Nayak"
date: "2025-04-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dir = "C:/cygwin64/home/vedan/Code/sta141b"
setwd(dir)

source(paste(dir, "/assignment2code.R", sep=''))
```

# Overview
The goal of this assignment is to create a dataframe containing information from server alert data. 

The final data table should include 23 columns:

- Title of Event
- Classification
- Priority
- Snort ID
- Month
- Day
- Timed
- IP address + Port
    - 4 columns total, 2 for the source and 2 for the destination
- The protocol
- TTL
- TOS 
- ID
- IpLen
- DgmLen
- Extra (the text after the DgmLen in line 4)
- Seq
- Ack
- Win 
- TcpLen
- TCP flag
- Extra lines
- Filename

# Functions

To more easily organize the information, created functions to extract the information from each line.

## getblocks()

This function looks for empty lines, does a cumsum on the true/false values, and splits the text based on those values. 
I noticed that I was getting an extra empty block in my result. I deleted the last block if the last value has a "TRUE" when looking for empty lines. 

## extractLine1()

1. Extract the line n elements from all the blocks. 
2. Use regular expressions and gsub() to extract the desired information. 
- I tested the regular expression incrementally to develop it 
- I saw that each line started with [**]
- The *Snort ID* will always be [integer:integer:integer]
- The *title* will be a mix of letters and spaces and will follow the Snort ID
3. Split the resulting vector elements by the designated space.
4. Convert the resulting vector into a matrix (using do.call() and rbind()) then dataframe.
5. Rename the column names to something more meaningful. 

## extractLine2()

1. Extract the line n elements from all the blocks. 
2. String split on *[* and *]* to break apart the *classification* and *proximity*. 
3. Remove empty strings
4. The 1st term will always be the*classification* and the second term will always be *proximity*. I used this to extrac the respective terms. 
5. Create a dataframe and rename the columns.

## extractLine3()

1. Extract the line n elements from all the blocks. 
2. Use regular expressions and gsub() to extract the desired information. 
- *Day/Month*: After extracting the day/month term (always integer/integer), string split based on the "/". I then added the resulting value into a matrix then dataframe. 
- *Time*: Integers following the day/month and that follow integer(2x):integer(2x):integer(2x).integer(some number). After getting the value I made another dataframe. 
- *IP & Port*: I extacted both IP information into one vector. I found the IP values using the pattern I built followed by a (wildcard) -> (wildcard). The source always preceded the destination, so when splitting along the ";", I extracted the first term for the source and the second for the destination from the resulting vector. I then split the result across the ":" and used the *makeNaLine3()* to fill in missing port values with the character "NA". I made a third dataframe with this information.
4. I then bound the resulting dataframes using cbind().

## extractLine4()

1. Extract the line n elements from all the blocks. 
2. Use regular expressions and gsub() to extract the desired information. 
- I saw that all the line4s follow the same order of values, so I made one regular expression for the 'Protocol', 'TTL', 'TOC', 'ID', 'IpLen', 'DgmLen', & 'Extra' values, respectively. 
- *Protocol*: letters(some number)
- *TTL*: integers(some number)
- *ID*: integers(some number)letters(somenumber)integers(some number)
- *IpLen*: integers(some number)
- *DgmLen*: integers(some number)
- *Extra*: wildcard(whatever is left)
3. Split the resulting vector elements by the designated space.
- For *blocks with no extra values* I used *makeNaLine4()* to fill the missing value with the charecter "NA". 
4. Convert the resulting vector into a matrix (using do.call() and rbind()) then dataframe.
5. Rename the column names to something more meaningful. 

## extractLine5()

1. Extract the line n elements from all the blocks. 
2. Use regular expressions and gsub() to extract the desired information. 
- I saw that all *the lines with a line5* follow the same pattern 'TCP_flag', "Seq", 'Ack', 'Window', & 'TCP_len' values, respectively. I used a wildcard to find the respective values and get a vector with the respective values. 
3. Split the resulting vector elements by the designated space.
- For the *blocks with a line5*, the workflow is similar to before. 
- For the *blocks with no line 5*, I used the *makeNaLine5()* to replace missing values with the character "NA".
4. Convert the resulting vector into a matrix (using do.call() and rbind()) then dataframe.
5. Rename the column names to something more meaningful.

## extractExtraLines()

1. Create a list of blocks without the first 5 lines.
2. Collapse the remaining lines using paste and separate the lines with a newline character. 
3. Turn the vector into a dataframe

## makeDF()

1. Read in the lines and use *getBlocks()* to extract the blocks. 
2. Pass the blocks to each of the *extract___()* functions to get 6 dataframes (one for each line/gtoup of lines)
3. Bind the dataframes together
4. Remove lines that contain NAs
- I kept getting an extra line at the end. Since I used the character "NA" to fill in missing values, using na.omit() over my dataframe was a suitable way to remove a row with nothing. 
5. I converted the month, day, proximity, TTL, ID, IpLen, and DgmLen into integers because those are the only numerical variables that will not contain missing values. 

## checkValid()

Use this function to determine if blocks containing a particular pattern are the same as blocks containing the pattern at the expected line. 

1. Make a vector with the block numbers containing the pattern and another containing the block numbers containing the pattern at the expected line. 
2. Output a TRUE/FALSE table comparing the two vectors. 

## whichNotValid()
Use this function to determine which blocks containing a particular pattern are the NOT the same as blocks containing the pattern at the expected line. 

1. Make a vector with the block numbers containing the pattern and another containing the block numbers containing the pattern at the expected line.
2. Output a vector containing lines that do not match up. 

# Validation

## Validating Regular Expressions Used

### Line 1

I first used *checkValid()* to confirm that the *SnortID* and *Title* were on line 1. 

```{r}
validLine1Pretty
```

Both the SnortID and Title are only found on line 1. 

Since both are only line 1, I selected one of the patterns as the comparison. Since they're all on line 4, I selected just one of the patterns as the comparison logical vector to confirm the regular expression is also true. 
I wanted to see if the lines containing regular expression I wrote is are the same as those that contain the Title, thereby proving the regular expression I wrote is correct. 

```{r}
validLine1RegEx
```

Since all values are true, my regular expression works.

### Line 2

I used *checkValid()* to confirm that the *Classification* and *Priority* were on line 2.

```{r}
validLine2Pretty
```

Since everything came back as true, both are found on line 2. 

I then wanted to confirm my assumption that classification comes before priority is always correct. I did this by validating the regular expression I used to extract data. Since they're all on line 2, I selected just one of the patterns as the comparison logical vector to confirm the regular expression is also true. 

```{r}
validLine2RegEx
```


### Line 3


I used *checkValid()* to confirm that the *Day/Month*, *Time*, and *IP* were on line 3.

```{r}
validLine3Pretty
```

Since everything came back as true, all values are found on line 3. 

### Line 4

I used *checkValid()* to confirm that the *Protocol*, *TTL*, *TOS*, *ID*, *IpLen* and *DgmLen* were on line 4.

```{r}
validLine4Pretty
```

Since all of them yielded true, these elements are only on line 4. Since they're all on line 4, I selected just one of the patterns as the comparison logical vector to confirm the regular expression is also true. 


```{r}
validLine4RegEx
```

The regular experession is true for all. 

### Line 5

I used *checkValid()* to confirm that the *Sequence*, *Acknowledgment*, *Window*, and *TcpLen* were on line 5. 

```{r}
validLine5Pretty
```

*Note: I manually changed the FALSE values for the Ack, Win, and TcpLen rows to "0" based on information I saw on the validLine5Els list.*

From this, we can see two things:

1. Elements "Ack:", "Win:", and "TcpLen:" are all on line 5 (since they all lines containing those values were the same as the lines containing those values on line 5)
2. Element "Seq:" was not only on line 5. 3262 lines contained "Seq:" which were not on line 5. I wanted to further investigate the blocks where "Seq:" was found on lines that were not line 5.

The lines that contained "Seq:" on lines other than line 5 also did not contain any of the other elements we need to extract (Acknowledgment, Window, and TcpLength). They contain other pieces of information instead, such as "Type". They are related to "Misc activity". While they contain an IP address, they do not contain a port for the IP value.

```{r}
noSeqSummaryTable
```



### Extra Lines



## Dataframe Validation


### Check the number of rows and columns

*Expected number*