---
title: "Reading Solar Data -- Part 1"
author: "Veda Nayak"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#I. Overview 

##What to Submit
 • A description of your approach and the important steps.
 • The file(s) with the R functions and the code to use those functions to read the data.
 • A description of how you verified the results were correct with numerical summaries, tables, plots that “prove” this is
 correct, along with a description of any manual verification you did.
 • The R code you used to verify the results

##Objective: 
For this assignment, you are to write functions that will read the data in the .clm (climate) file in each zip file into a data.frame. The resulting *data.frame should have 9 columns– the 6 columns in the data set and also the day, month and hour of each observation.*
Write functions (rather than one or more R commands) to read each of these tables, as you need to apply these to the contents of the 5 ZIP files. Also, you will most likely need to run the code multiple times to iteratively modify it and verify it is correct. Use functions rather than repeating the same code for each file and table

##Verification: Describe the approaches / processes taken to verify your findings

#II. Trial Run on a Small Dataset (three days)

I created a dataset taking lines 1 - 87 of the "USA_CA_Mammoth.Yosemite.AP.723894_TMYx.2009-2023.clm" file.

I treated the .clm file like a .txt file. 

```{r}
setwd("C:/cygwin64/home/vedan/Code/sta141b")

lines.list <- readLines("test.clm")

# Get the line number that starts each day
# https://www.spsanderson.com/steveondata/posts/2024-09-03/
matching_indices <- grep("* day  ", lines.list)
print(matching_indices)
print(lines.list[matching_indices])

test_frame <- as.data.frame(matrix(0, nrow = 24*length(matching_indices), ncol = 9))
colnames(test_frame) <- c('day', 'month', 'hour', 'diffuse_solar_horizontal', 'dry_bulb_temp', 'normal_solar_intensity', 'wind_speed', 'wind_direction', 'rel_humidity')

start = 1
for (i in 1:length(matching_indices)){
  
  # get the text contents for a particular day
  line_index = as.integer(matching_indices[i])
  day_info <- lines.list[line_index:(line_index+24)]
  day_split <- strsplit(day_info[1], split = " ")
  unlist(day_split)
  
  # day and month information
  day <- as.integer(day_split[[1]][4])
  month <- as.integer(day_split[[1]][7])
  
  # fill in the hour and remaining 6 columns
  if (i == 1){start = 1}
  else {start = (i - 1) * 24 + 1}
  
  stop = start + 24
  
  hour = 1
  
  # remove the first line
  line_index = 38
  index_plus_one = line_index + 1
  ending_index = line_index + 24
  data_info <- lines.list[index_plus_one:ending_index]
  
  line_index = 1
  for (j in start:stop){
    
    test_frame[j,1] <- day
    test_frame[j,2] <- month
   
    # add in the hours 
    test_frame[j,3] <- hour
    
    line = strsplit(data_info[line_index], split = ',')

    test_frame[j, 4] <- line[[1]][1]
    test_frame[j, 5] <- line[[1]][2]
    test_frame[j, 6] <- line[[1]][3]
    test_frame[j, 7] <- line[[1]][4]
    test_frame[j, 8] <- line[[1]][5]
    test_frame[j, 9] <- line[[1]][6]
  
    line_index = line_index + 1
    
    }
    
    hour = hour + 1
    
  }
  

```
I then wrapped these steps into a function.

```{r}

solar.data.frame <- function(lines.list) {
  

  # find the line number that starts each day
  
  matching_indices <- grep("* day  ", lines.list)
  print(matching_indices)
  print(lines.list[matching_indices])
  
  test_frame <- as.data.frame(matrix(0, nrow = 24*length(matching_indices), ncol = 9))
  colnames(test_frame) <- c('day', 'month', 'hour', 'diffuse_solar_horizontal', 'dry_bulb_temp', 'normal_solar_intensity', 'wind_speed', 'wind_direction', 'rel_humidity')
  
  start = 1
  for (i in 1:length(matching_indices)){
    
    # get the text contents for a particular day
    line_index = as.integer(matching_indices[i])
    day_info <- lines.list[line_index:(line_index+24)]
    day_split <- strsplit(day_info[1], split = " ")
    unlist(day_split)
    
    # day and month information
    day <- as.integer(day_split[[1]][4])
    month <- as.integer(day_split[[1]][7])
    
    # fill in the hour and remaining 6 columns
    if (i == 1){start = 1}
    else {start = (i - 1) * 24 + 1}
    
    stop = start + 24
    
    hour = 1
    
    # remove the first line
    line_index = 38
    index_plus_one = line_index + 1
    ending_index = line_index + 24
    data_info <- lines.list[index_plus_one:ending_index]
    
    line_index = 1
    for (j in start:stop){
      
      test_frame[j,1] <- day
      test_frame[j,2] <- month
     
      # add in the hours 
      test_frame[j,3] <- hour
      
      line = strsplit(data_info[line_index], split = ',')
  
      test_frame[j, 4] <- line[[1]][1]
      test_frame[j, 5] <- line[[1]][2]
      test_frame[j, 6] <- line[[1]][3]
      test_frame[j, 7] <- line[[1]][4]
      test_frame[j, 8] <- line[[1]][5]
      test_frame[j, 9] <- line[[1]][6]
    
      line_index = line_index + 1
      
      }
      
      hour = hour + 1
      
    }
  return(test_frame)
}

data <- readLines(unz("USA_CA_Mammoth.Yosemite.AP.723894_TMYx.2009-2023.zip", "USA_CA_Mammoth.Yosemite.AP.723894_TMYx.2009-2023.clm"))

df <- solar.data.frame("text.clm")
```


